---
title: "Zomato Bangalore Restaurant Analysis"
author: "Rachit Malik"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(janitor)
library(skimr)
library(stringr)
```

# Introduction

This notebook documents an end-to-end **Zomato Bangalore Restaurant Analysis**. The goal is to understand what drives restaurant success using data-driven insights and to prepare clean datasets for **Tableau / Power BI dashboards**.

The workflow follows a structured analytics process:

**Prepare → Process → Analyze → Export**

---

#  Data Loading & Initial Inspection

```{r load-data}
zomato_data <- read.csv("C:/Users/a2z/Downloads/zomato.csv/zomato.csv")

glimpse(zomato_data)
skim_without_charts(zomato_data)
```

---

# Data Health Check

```{r data-health}
# Missing values per column
colSums(is.na(zomato_data))

# Duplicate rows
sum(duplicated(zomato_data))

# Inspect messy columns
head(unique(zomato_data$rate), 20)
head(unique(zomato_data$approx_cost.for.two.people.), 20)
unique(zomato_data$online_order)
unique(zomato_data$book_table)
```

**Key Issues Identified**
- Ratings include text values like `NEW` and `-`
- Cost column contains commas
- Online order & table booking stored as text

---

# Data Cleaning & Feature Engineering

```{r data-cleaning}
# Rename and clean column names
zomato_v1 <- zomato_data %>% 
  rename(cost = approx_cost.for.two.people.,
         type = listed_in.type.,
         city = listed_in.city.) %>% 
  clean_names()

# Fix data types and values
zomato_v2 <- zomato_v1 %>% 
  mutate(
    rate = str_remove(rate, "/5"),
    rate = ifelse(rate %in% c("NEW", "-"), NA, rate),
    rate = as.numeric(rate),
    cost = str_remove(cost, ","),
    cost = as.numeric(cost)
  )

# Create binary flags for operations
zomato_final <- zomato_v2 %>% 
  mutate(
    has_online = ifelse(online_order == "Yes", 1, 0),
    has_table_booking = ifelse(book_table == "Yes", 1, 0)
  ) %>% 
  drop_na(cost)

glimpse(zomato_final)
summary(zomato_final)
```

---

# Analysis Phase

## Analysis 1: Location Saturation & Opportunity

```{r location-analysis}
loc_analysis <- zomato_final %>% 
  group_by(location) %>% 
  summarise(
    restaurant_count = n(),
    avg_rating = mean(rate, na.rm = TRUE),
    avg_cost = mean(cost, na.rm = TRUE)
  ) %>%
  filter(restaurant_count > 50) %>%
  arrange(desc(restaurant_count))

head(loc_analysis, 10)
head(arrange(loc_analysis, desc(avg_rating)), 10)
```

---

## Analysis 2: Impact of Online Ordering & Table Booking

```{r tech-impact}
tech_impact <- zomato_final %>% 
  group_by(has_online, has_table_booking) %>% 
  summarise(
    count = n(),
    avg_rating = mean(rate, na.rm = TRUE),
    avg_votes = mean(votes, na.rm = TRUE),
    .groups = "drop"
  )

tech_impact
```

---

## Analysis 3: Price vs Rating Relationship

```{r price-rating}
ggplot(zomato_final, aes(x = cost, y = rate)) +
  geom_point(alpha = 0.3, color = "darkblue") +
  geom_smooth(method = "lm", color = "red") +
  labs(
    title = "Does Cost Affect Rating?",
    subtitle = "Price vs Quality Relationship",
    x = "Cost for Two (INR)",
    y = "Rating (1–5)"
  ) +
  theme_minimal()
```

---

## Analysis 4: Top Tier vs Average Restaurants (Spider Chart Prep)

```{r radar-prep}
radar_data <- zomato_final %>%
  mutate(segment = ifelse(rate >= 4.5, "Top Tier (Top 1%)", "Average Restaurant")) %>%
  group_by(segment) %>%
  summarise(
    Avg_Rating = mean(rate, na.rm = TRUE),
    Avg_Cost = mean(cost, na.rm = TRUE),
    Avg_Votes = mean(votes, na.rm = TRUE),
    Online_order = mean(has_online, na.rm = TRUE) * 100,
    Booking_Available = mean(has_table_booking, na.rm = TRUE) * 100,
    .groups = "drop"
  )

radar_clean <- radar_data %>% drop_na()
radar_clean
```

---

# 🍽️ Cuisine & Dish Analysis

## Top Cuisines

```{r top-cuisines}
top_cuisines <- zomato_final %>%
  drop_na(cuisines) %>%
  separate_rows(cuisines, sep = ",") %>%
  mutate(cuisines = str_trim(cuisines)) %>%
  count(cuisines, sort = TRUE) %>%
  head(30)

top_cuisines
```

## Most Loved Dishes (Top Restaurants)

```{r top-dishes}
top_dishes <- zomato_final %>%
  filter(rate >= 4.0) %>%
  drop_na(dish_liked) %>%
  separate_rows(dish_liked, sep = ",") %>%
  mutate(dish_liked = str_to_lower(str_trim(dish_liked))) %>%
  filter(!dish_liked %in% c("friendly staff", "good ambience", "service", "cocktails")) %>%
  count(dish_liked, sort = TRUE) %>%
  head(50)

top_dishes
```

---

# Export Clean Data for BI Tools

```{r export}
write.csv(zomato_final, "C:/Users/a2z/Downloads/zomato_clean_data.csv", row.names = FALSE)
write.csv(radar_clean, "C:/Users/a2z/Downloads/zomato_radar_profile.csv", row.names = FALSE)
write.csv(top_cuisines, "C:/Users/a2z/Downloads/zomato_top_cuisines.csv", row.names = FALSE)
write.csv(top_dishes, "C:/Users/a2z/Downloads/zomato_top_liked_dishes.csv", row.names = FALSE)

getwd()
```

---

# Conclusion

This analysis reveals that **location saturation**, **operational technology (online ordering & booking)**, and **pricing balance** are strong indicators of restaurant success. The cleaned datasets are now ready for **Tableau / Power BI dashboards and strategic recommendations**.

```{r session-info, echo=FALSE}
sessionInfo()